How to compute the fusion product of MV cycles in type A

### 2021-02-16 draft.tex
more minor edits

The Beilinson--Drinfeld Grassmannian, essential to \cite{mirkovic2007geometric}, is a space which is used to define a fusion (aka convolution) product on sheaves (representations) and MV cycles (bases). 
% 
This product structure makes the vector space spanned by the MV cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2020bases}). % conjectured by JEA and proved by BKK
%

@@

Fix $G = \GL_m$ and let $T\subset G$ be the maximal torus of diagonal matrices. 
% are there other possibilities?? :S 
Write $P$ for the character lattice of $T$. Identify $P\cong \ZZ^m$ so that the set of dominant weights $P_+$ is described/given by the set of partitions having at most $m$ nonzero parts. Set $\cO = \CC\xt$ and $\cK = \CC\xT$. We will also need the notation $\cO_x$ for the completion of the local ring $\CC[t]$ at $x\in\spec\CC[t]$ and $\cK_x$ for its fraction field $\CC\xT[t-x]$. 
% 
Note that (by choosing the local coordinate $t-x$ on $\spec\CC[t]$) we can identify $\cO_x \cong \cO$ and $\cK_x\cong\cK$.

@@

\begin{itemize}
    \item The ordinary affine Grassmannian $\Gr = G(\cK)/G(\cO)$ % switched out ordinary Gr notation for curly Gr 
    \item The Beauville--Laszlo Grassmannian $\Grbd[x] = G(\cK_x)/G(\cO_x)$ % TODO: in terms of bundles it is $\{(P,\sigma) : P is a PGB over A and \sigma is a trivialization of P on A - x\}$
    \item The (two point) Beilinson--Drinfeld Grassmannian $\pi : \Grbd\to \AA$ (with one point fixed at 0) % switched out general curve $C$ for affine line $\AA$; TODO: in terms of bundles it is $\{(P,\sigma,x):agaiin P is a PGB on A, x \in A, and now \sigma is a trivialization away from 0 _and_ x \}
    \item Partitions $\mu_i\le\lambda_i$ of $N_i$ ($i=1,2$) and $\mu = \sum \mu_i \le \lambda = \sum\lambda_i$ of $N = \sum N_i$
    \item The slices $\Gr_\mu$ and $\cW_{\mu_1,\mu_2}$ to the orbits $\Gr^\lambda$ and $(\Grbd[2])^{\lambda_1,\lambda_2}$
    \item The nilpotent and semi-nilpotent cones $\cN$ and $\cN_s$ (of matrices with eigenvalues 0 and 0 or $s\ne 0$)
    \item The slices $\TT_\mu$ and $\TT_{\mu_1,\mu_2}$ to the orbits $\OO_\lambda$ and $\OO_{\lambda_1,\lambda_2}$ 
\end{itemize}

New (?) definitions among these are as follows.
%

%The [family of] slices [with $s$-fibre?]
%\begin{equation}
%    \cW_{\mu_1,\mu_2} = G_1 [t^{-1},(t-s)^{-1}]L_{\mu_1,\mu_2}
%\end{equation}

We define
$$L_\mu = t^\mu, L_{\mu_1, \mu_2} = t^{\mu_1} (t-s)^{\mu_2} \in G((t^{-1}))/G[t]$$

\begin{equation}
    \cW_\mu = G_1[[t^{-1}]]L_\mu \subset G((t^{-1}))/G[t]
\end{equation}
a subscheme of the thick affine Grassmannian

Note that $ L_{\mu_1, \mu_2} = (t-s)^{\mu_2} t^{-\mu_2} L_{\mu} \in \cW_\mu$
and that  $L_{\mu_1,\mu_2}\in\Grbd[2]$ is a $\CC[t]$-lattice in $\CC(t)^m$ that specializes to a $\CC\xt$-lattice in $\CC\xT^m$ away from $t = 0$ and away from $t = s$; i.e.\
\begin{equation}
    \label{eq:defWBD}
    \begin{split}
        %L_{\mu_1,\mu_2}\otimes \CC[(t-s)^{-1}] =
        L_{\mu_1,\mu_2}\otimes \CC\xt &= L_{\mu_2} \otimes \CC\xt \text{ and } \\
        %L_{\mu_1,\mu_2}\otimes\CC[t^{-1}] =
        L_{\mu_1,\mu_2}\otimes \CC\xt[t-s] &= L_{\mu_1} \otimes \CC\xt[t-s]
    \end{split}
\end{equation}
where $L_{\mu_i}$ denotes the point $t^{\mu_i}G(\cO)\in\Gr$. 
%

The family of semi-infinite orbits 
\begin{equation}
    S_{\mu_1,\mu_2} = N_-(\CC((t^{-1}))) L_{\mu_1,\mu_2}\,. 
\end{equation}

The orbit of $\CC[t]$-lattices in $\CC(t)^m$ whose elements specialize again to $\CC\xt$-lattices in $\CC\xT^m$ away from $t = 0$ and away from $t = s$ as follows
\begin{equation}
    \label{eq:defGrBDlambda}
    \begin{split}
        (\Grbd[2])^{\lambda_1,\lambda_2} = \{L \in\Grbd[2] : 
        %L\otimes\CC[t^{-1}] 
        L \otimes_{\CC[t]} \CC\xt[t-s] 
        &\in \Gr^{\lambda_2} \text{ and } \\
        % L\otimes\CC[(s-t)^{-1}] 
        L\otimes_{\CC[t]} \CC\xt &\in \Gr^{\lambda_1} \text{ and } \\ 
        L \otimes_{\CC[t]} \CC\xt[t-a] = L_0 \text{ for all $ a \ne 0, s$} \}\,.
    \end{split}
\end{equation}

\jcom{It is important to add the condition that the lattices are trivial at $ a $.}
% 
\acom{These defining conditions are again just telling us that in the fibre over the fixed point $(0,s)\in C^{(2)}$ this set is pairs of lattices; note that we invert the indeterminates $t$ and $(t-s)$ \textit{after} specializing $(0,s)$. Could we give a more explicit characterization like $t^{N_1}(t-s)^{N_2} L_0 \subseteq L\subseteq t^{-N_1}(t-s)^{-N_2}$?}

@@

In Anderson and Kogan conjectured in \cite{anderson2006algebra} \acom{or earlier?} that those MV polynomials which are cluster monomials for a Fomin--Zelevinsky cluster algebra structure on $\CC[N]$ are naturally expressible as determinants\dots
and they conjecture a formula for many of them.

It's not clear how this work helps/relates to AK/their conjectures.

The generalized MVy is interesting in its own right.

@@

\begin{quotation}
However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}

@@

\begin{theorem}
    Let $\lambda_i\ge\mu_i$ be dominant ($i=1,2$), $\mu = \mu_1 +\mu_2$, and $\lambda =\lambda_1+\lambda_2$. 
    % I have an irrational dislike of ending sentences with subscripts --- can we fix this?
    There is an isomorphism 
    \begin{equation}
        \overline{\OO_{\lambda_1,\lambda_2}}\cap\TT_{\mu_1,\mu_2} \to \overline{(\Grbd[2])^{\lambda_1,\lambda_2}}\cap \cW_{\mu_1,\mu_2}
    \end{equation}
    got by taking a $\mu\times\mu$ block matrix $A$ in the $s$-fibre $\overline{\OO_{\lambda_1,\lambda_2}^s}\cap\TT_{\mu_1,\mu_2}^s$ on the left to the representative of the $s$-fibre on the right defined by  
    \begin{equation}
        \begin{split}
            g &= t^{\mu_1} (t-s)^{\mu_2} + a(t) \\
            a_{ij}(t) &= - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1}
        \end{split}
    \end{equation}
    where $A^k_{ji}$ is the $k$th entry from the left of the last row of the $\mu_j\times\mu_i$ block of $A$. 
\end{theorem}

>>>>

The Beilinson--Drinfeld Grassmannian, essential to \cite{mirkovic2007geometric}, is a space which is used to define a fusion (aka convolution) product on sheaves (representations) and MV cycles (bases). 
% 
This product structure makes the vector space spanned by the MV cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2020bases}). % conjectured by JEA and proved by BKK
% 
\begin{quotation}
    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}

@@

Fix $G = \GL_m$ and let $T\subset G$ be the maximal torus of diagonal matrices. 
% are there other possibilities?? :S 
Write $P$ for the character lattice of $T$. Identify $P\cong \ZZ^m$ so that the set of dominant weights $P_+$ is described/given by the set of partitions having at most $m$ nonzero parts. Set $\cO = \CC\xt$ and $\cK = \CC\xT$. We will also need the notation $\cO_s$ for the completion of the local ring $\CC[t]$ at $s\in\spec\CC[t]$ and $\cK_s$ for its fraction field $\CC\xT[t-s]$. 
% 
Note that (by choosing the local coordinate $t-s$ on $\spec\CC[t]$) we can identify $\cO_s \cong \cO$ and $\cK_s\cong\cK$.

@@

\begin{itemize}
    \item The ordinary affine Grassmannian $\Gr = G(\cK)/G(\cO)$ % switched out ordinary Gr notation for curly Gr 
    \item The Beauville--Laszlo Grassmannian $\Grbd[s] = G(\cK_s)/G(\cO_s)$ % TODO: in terms of bundles it is $\{(P,\sigma) : P is a PGB over A and \sigma is a trivialization of P on A - x\}$
    \item The (two point) Beilinson--Drinfeld Grassmannian $\pi : \Grbd\to \AA$ (with one point fixed at 0; and the second point $s$ varying away from $s = 0$) % switched out general curve $C$ for affine line $\AA$; TODO: in terms of bundles it is $\{(P,\sigma,x):agaiin P is a PGB on A, x \in A, and now \sigma is a trivialization away from 0 _and_ x \}
    \item Weights $\mu_i\le\lambda_i$ of size $N_i$ ($i=1,2$) 
    \item Their sums $\mu = \sum \mu_i \le \lambda = \sum\lambda_i$ of size $N = \sum N_i$
    \item The orbit $\Gr^\lambda = G(\cO) L_\lambda$ \acom{and its closure?}
    \item The orbit $\Gr_\AA^{\lambda_1,\lambda_2}\to\AA$ whose generic fibre is $\Gr^{\lambda_1} \times \Gr^{\lambda_2}$ and whose zero fibre is $\Gr^{\lambda}$ \acom{and closures?}
    \item The slice $\Gr_\mu = G_1[t^{-1}]L_\mu$ to the orbit $\Gr^\lambda$
    \item The slice $\cW_\mu = G_1\xT[t^{-1}] L_\mu$ to the orbit $\Gr_\AA^{\lambda_1,\lambda_2}$ \acom{is not presented as a family; why need define it in thick Gr?}% $(\Grbd[2])^{\lambda_1,\lambda_2}$; replaced \cW_{\mu_1,\mu_2} = by Joel's not'n & def, seemingly independent of the decomposition of \mu; modified Joel's double bracket --- still confused over where W_\mu ought to live 
    \item The nilpotent and semi-nilpotent cones $\cN$ and $\cN_\AA$ (of matrices with eigenvalues in $\{0\}$ and $\{0,s\} \in \AA\times\AA$ resp.) 
    \item The slice $\TT_\mu$ to the orbits $\OO_\lambda$ 
    \item The family of slices $\TT_{\mu_1,\mu_2}^{\AA}$ to the family of orbits $\OO^{\AA}_{\lambda_1,\lambda_2}$ 
\end{itemize}

% New (?) definitions among these are as follows.
% 
%The [family of] slices [with $s$-fibre?]
%\begin{equation}
%    \cW_{\mu_1,\mu_2} = G_1 [t^{-1},(t-s)^{-1}]L_{\mu_1,\mu_2}
%\end{equation}

Recall (from \cite{kamnitzer2014yangians}) that there is an embedding of the usual ``thin'' affine Grassmannian $\Gr $ as we have defined it into the ``thick'' affine Grassmannian. \acom{In \cite{kamnitzer2014yangians} the thick Gr is said to be necessary for the noncommutative algebras considered. What is our reason for preferring it to the thin Gr? Can we replace thick by thin?}
\[
\Gr \cong G[t,t^{-1}]/G[t]\hookrightarrow G\xT[t^{-1}]/G[t]\,.
\]
Let $L_\mu = t^\mu G(\cO)$ and $L_{\mu_1, \mu_2} = t^{\mu_1} (t-s)^{\mu_2} G(\cO)$. \acom{or $ t^{\mu_1} (t-s)^{\mu_2} G[t]$ in thick?} % \in G((t^{-1}))/G[t]

% \begin{equation}
%     \cW_\mu = G_1[[t^{-1}]]L_\mu \subset G((t^{-1}))/G[t]
% \end{equation}
% a subscheme of the thick affine Grassmannian

Note that $ L_{\mu_1, \mu_2} = (t-s)^{\mu_2} t^{-\mu_2} L_{\mu} \in \cW_\mu$ \acom{thick or thin}
and that  $L_{\mu_1,\mu_2}\in\Grbd$ \acom{should say $\Grbd[s]$ as it's really a lattice in a given fibre being characterized?} is a $\CC[t]$-lattice in $\CC(t)^m$ that specializes to a $\CC\xt$-lattice in $\CC\xT^m$ away from $t = 0$ and away from $t = s$ \acom{thick only} i.e.\
\begin{equation}
    \label{eq:defWBD}
    % \begin{split}
        %L_{\mu_1,\mu_2}\otimes \CC[(t-s)^{-1}] =
        L_{\mu_1,\mu_2}\otimes \CC\xt = L_{\mu_2} \otimes \CC\xt \qquad
        L_{\mu_1,\mu_2}\otimes \CC\xt[t-s] = L_{\mu_1} \otimes \CC\xt[t-s]
    % \end{split}
\end{equation}
where 
% $L_{\mu_i}$ denotes the point $t^{\mu_i}G(\cO)\in\Gr$ and 
the tensor product is taken over $\CC[t]$. 
%

The family of semi-infinite orbits 
\begin{equation}
    S_{\mu_1,\mu_2} = N_-(\CC\xT[t^{-1}]) L_{\mu_1,\mu_2}\,. 
\end{equation}
\acom{1. Introduce $N_\pm$. 2. In \cite{kamnitzer2014yangians} the global version of $S^\mu_-$ is the subvariety $S^\mu_{-,\AA}\subset\Grbd$ whose generic fibre is $\bigcup_{\mu_1 + \mu_2 = \mu} S^{\mu_1}_-\times S^{\mu_2}_-$ and whose zero fibre is $S^\mu_-$ which don't look to be the same thing.}

The family \acom{orbit? orbits? family of orbits?} of $\CC[t]$-lattices in $\CC(t)^m$ whose elements specialize again to $\CC\xt$-lattices in $\CC\xT^m$ away from $t = 0$ and away from $t = s$ as described more explicitly as follows
\begin{equation}
    \label{eq:defGrBDlambda}
    \begin{split}
        % (\Grbd[2])^{\lambda_1,\lambda_2} = 
        \Gr_{\AA}^{\lambda_1,\lambda_2} = 
        \{L \in\Grbd : 
        %L\otimes\CC[t^{-1}] 
        L \otimes_{\CC[t]} \CC\xt[t-s] 
        &\in \Gr^{\lambda_2} \text{ and } \\
        % L\otimes\CC[(s-t)^{-1}] 
        L\otimes_{\CC[t]} \CC\xt &\in \Gr^{\lambda_1} \text{ and } \\ 
        L \otimes_{\CC[t]} \CC\xt[t-a] &= L_0 \text{ for all $ a \ne 0, s$} \}\,.
    \end{split}
\end{equation}

\jcom{It is important to add the condition that the lattices are trivial at $ a $.} \acom{$a\ne 0,s$}
% 
% \acom{These defining conditions are again just telling us that in the fibre over the fixed point $(0,s)\in C^{(2)}$ this set is pairs of lattices; note that we invert the indeterminates $t$ and $(t-s)$ \textit{after} specializing $(0,s)$. Could we give a more explicit characterization like $t^{N_1}(t-s)^{N_2} L_0 \subseteq L\subseteq t^{-N_1}(t-s)^{-N_2}$?}

@@

In Anderson and Kogan conjectured in \cite{anderson2006algebra} \acom{or earlier?} that those MV polynomials which are cluster monomials for a Fomin--Zelevinsky cluster algebra structure on $\CC[N]$ are naturally expressible as determinants\dots
% and they conjectured a formula for many of them.

It's not clear how this work helps with/relates to AK/their conjectures.

The generalized MVy is interesting in its own right. \acom{is it?}

@@

Interesting combinatorics? Witnessing a product on SSYT.

@@

\begin{theorem}
    Let $\lambda_i\ge\mu_i$ be dominant ($i=1,2$), $\mu = \mu_1 +\mu_2$, and $\lambda =\lambda_1+\lambda_2$. 
    % I have an irrational dislike of ending sentences with subscripts --- can we fix this?
    There is an isomorphism 
    \begin{equation}
        \overline{\OO}_{\lambda_1,\lambda_2}\cap\TT_{\mu_1,\mu_2} \to \overline\Gr_\AA^{\lambda_1,\lambda_2}\cap \cW_{\mu_1,\mu_2}
    \end{equation}
    got by taking a $\mu\times\mu$ block matrix $A$ in the $s$-fibre $\overline{\OO_{\lambda_1,\lambda_2}^s}\cap\TT_{\mu_1,\mu_2}^s$ on the left to the representative of the $s$-fibre on the right defined by  
    \begin{equation}
        \begin{split}
            g &= t^{\mu_1} (t-s)^{\mu_2} + a(t) \\
            a_{ij}(t) &= - \sum_{k=1}^{\mu_i} A^k_{ji} t^{k-1}
        \end{split}
    \end{equation}
    where $A^k_{ji}$ is the $k$th entry from the left of the last row of the $\mu_j\times\mu_i$ block of $A$. 
\end{theorem}
